Whittaker Models for Real Groups

### 2024-06-12 whittaker.tex
Fairly serious round of changes
Fixed missing *'s and signs in many places, added some missing proofs
New Fresse/Mehdi reference

\subsubsection*{Nilpotent orbits}

Let $\N$ be the set of nilpotent elements of $\g$.
This is the closure of the principal nilpotent orbit $\Op$.\footnote{(AA 2024-V-1) Todo: give nice references...}

@@

Suppose $\pi$ is an irreducible $(\g,K)$-module.
Let $I_\pi$ be the annihlator of $\pi$ in the universal enveloping algebra.
Let $\AVann(\pi)$ be the associated variety of $I_\pi$.
This is the closure of a single complex nilpotent orbit $\O$.
The associated variety of $\pi$, denoted $\AV(\pi)$, is a closed, $K$-invariant set in $\s$, so it is a finite union of
nilpotent $K$-orbits on $\s$.\footnote{(AA 2024-V-1) Todo: give nice references, Vogan's Bowdoin paper and others...}.

@@

If we do not start with a $(\g, K)$-module but with a Hilbert space representation~$\pi$ of $G(\R)$,  we define $\AV(\pi)$ to be the associated variety 
of its underlying $(\g,K)$-module.

@@

If $\pi$ is an irreducible $(\g,K)$-module we define $\WF(\pi)$ to be the wave-front set of 
the Hilbert space globalization of $\pi$.

@@

We write $\Wh(\pi)$ for the set of Whittaker data~$\w$ such that If $\pi$ is an irreducible $(\g,K)$-module we define $\Wh(\pi)$ to be the set of 
the Hilbert space globalization of $\pi$.

@@

\begin{proposition} \label{invariants_ds}
  Suppose $\pi$ is a generic discrete series representation of $G(\R)$. Then:
\begin{enumerate}[(a)]
\item $\Wh(\pi)$ is a single Whittaker datum;
\item $\AV(\pi)$ is the closure of a single $K$-orbit on~$\mathcal{O}_p \cap \mathfrak{s}$;
\item  $\WF(\pi)$ is the closure of a single $G(\R)$-orbit on $\mathcal{O}_p \cap  i \g(\R)^*$.
\end{enumerate}
\end{proposition}

\begin{proof} For (a), by \cite{vogan-gelfand-kirillov} and \cite{kostant_whittaker}
a representation $\pi$ is large if and only if it admits a Whittaker model for some Whittaker datum,
and by \cite[Lemma 14.14]{abv} a  large discrete series representation admits a unique Whittaker datum.

Part~(b) is one of the main results of~\cite{vogan_bowdoin}: see \cite[Theorem ?]{vogan_bowdoin}.\footnote{\textcolor{red}{(AA, 2024-V-1) : TODO -- find the right references, I tried but quickly got lost...}}

@@

Because of this result we abuse notation slightly as follows. If~$\pi$ is a generic discrete series representation of~$G(\R)$, we write $AV(\pi)=\O$ to indicate that $AV(\pi)$ is the closure of the $K$-orbit $\O\subset \s$.
Similarly we write $\WF(\pi)=\O$ to indicate that $\WF(\pi)$ is the closure of the $G(\R)$-orbit $\O\subset i\g(\R)^*$.

@@

\begin{lemma}
  \label{l:large}
  The following conditions are equivalent.
  \begin{enumerate}
    \item $\pi$ is large;
\item $\AVann(\pi)=\N=\overline{\Op}$;
\item $\dim(\AVann(\pi))=N$;
  \item $\dim(\AV(\pi))=N/2$;
  \item $\GK(\pi)=N/2$;
    \item $\pi$ is generic.

@@

Given a generic discrete series representation~$\pi$ we consider the invariants of~$\pi$ defined by Proposition~\ref{invariants_ds}: a Whittaker datum  $\Wh(\pi)$, a $K$-orbit $\AV(\pi)$ in $\mathcal{O}_p \cap \mathfrak{s}$, and a $G(\R)$-orbit $\WF(\pi)$ in $(\mathcal{O}_p \cap  i \g(\R))$.

\begin{theorem} \label{th:main} Suppose~$\Pi$ is an $L$-packet of discrete series for~$G(\R)$. 
The maps $\pi \mapsto \Wh(\pi)$, $\pi \mapsto \AV(\pi)$ and $\pi\mapsto \WF(\pi)$ induce bijections between:
\begin{enumerate}
\item[(1)] The set~$\Pi_{\mathrm{gen}}$ of generic discrete series representations in~$\Pi$ ;
\item[(2)] The set of Whittaker data for $G(\R)$ ;
\item[(3)] The set of~$K$-orbits on $\mathcal{O}_p \cap \s$.
\item[(4)] The set of~$G(\R)$-orbits on $\mathcal{O}_p \cap  i \g(\R)^*$.
\end{enumerate}
\end{theorem}

Before embarking on the proof, let us point out that the maps in the theorem are defined using rather deep results of representation theory. However, we shall prove in Section~\ref{sec:explicit} that all bijections $(i) \leftrightarrow (j)$, for $i,j \in \{(1), \dots, (4)\}$, can be spelled out in elementary terms. Furthermore we shall prove, using elementary arguments, that (3) $\leftrightarrow$ (4) coincides with the Kostant--Sekiguchi correspondence.

@@

Note that $Q_\sigma(G)$ is a group of automorphisms of $G(\R)$. It therefore acts on representations of $G(\R)$. 
Similarly $Q_\theta(G)$ is a group of automorphisms of $(\g,K)$, and acts on $(\g,K)$-modules.\footnote{AA 30-IV-2024: shouldn't this be:  $\Gad^\sigma$ and $\Gad^\theta$  act on $G(\R)$-representations or $(\g,K)$-modules respectively, and this induces actions of $Q_{\sigma}(G)$ and $Q_{\theta}(G)$ on equivalence classes...?}
These two actions are intertwined by the isomorphism of Lemma~\ref{lem:iso_q}:

@@

To define the action of~$Q_{\sigma}(G)$ on~(4), first observe that~$\Gad$ acts on~$\g$ by the adjoint action. This action preserves~$\Op$, and its restriction to~$\Gad(\R)$ preserves~$\Op\cap i\g(\R)$. Since elements on a given~$G(\R)_{\mathrm{ad}}$-orbit are in the same~$G(\R)$-orbit, this induces an action of $Q_{\sigma}(G)=\Gad(\R)/G(\R)_{\mathrm{ad}}$ on~$(\Op\cap i\g(\R))/G(\R)$.

\begin{proposition}\label{prop:action_on_real_orbits}
\begin{enumerate} 
\item The action of~$Q_{\sigma}(G)$ on $(\Op\cap i\g(\R))/G(\R)$ is simply transitive.
\item The map $\pi \mapsto \WF(\pi)$ is $Q_{\sigma}(G)$-equivariant.
\end{enumerate}
\end{proposition}

\begin{proof}
The proof of~(1) is a standard argument in group cohomology. Suppose $\omega,\omega'$ are $G(\R)$ orbits in $\Op\cap i\g(\R)$,
and choose $X\in \omega,X'\in\omega'$. Then there exists $g\in G$ such that $\Ad(g)(X)=X'$. Since $X\in \g(\R)$ the condition $X'\in i\g(\R)$ 
is $g\inv \sigma(g)\in \Stab_G(X)$.  Since $X$ is principal $\Stab_G(X)=ZU$ where $U$ is a unipotent group. 
A standard fact is that $H^1(\sigma, U)=1$ \cite[Chap.~III, Proposition~6]{Serre_Galois}, and from this we see that, after multiplying $g$ on the right by an element of $u$, we can assume $g\inv \sigma(g)\in Z$. This is equivalent to: the image of $g$ in $\Gad$ is in $\Gad(\R)$. On the other hand $\omega=\omega'$ 
if and only if $X'=\Ad(g)X$ for some $g\in G(\R)$. This competes the proof of (1).

@@

On the other hand $Q(G)$, realized as $Q_\theta(G)$ is a group of automorphisms of $(\g,K)$.
It induces actions on $\s$, $\Op$ and $\Op\cap\s$, and on $(\g,K)$-modules.

\begin{proposition}\label{prop:action_on_K_orbits}
\begin{enumerate} 
\item The action of~$Q(G)$ on $(\Op \cap \s)/K$ is simply transitive.
\item The map $\pi \mapsto \AV(\pi)$ is $Q(G)$-equivariant.
\end{enumerate}
\end{proposition}

@@

First we describe the set of Whittaker data more explicitly.
Let $B$ be a Borel subgroup of $G$. Let $N$ be the nilradical of $B$, and $\overline N$ the opposite
nilpotent subgroup. Let $\n,\overline\n$ be the Lie algebras of $N,\overline N$.
If $B$ is defined over $\R$ we consider $N(\R),\n(\R), \overline\n(\R)$, etc.
Let $\kappa(\,,\,)$  be the Killing form. 
For $X\in i\overline \n(\R)$ define a unitary character $\psi_X$ of $N(\R)$ by:
$$
\psi_X(e^Y)=e^{2\pi \kappa(X,Y)}\quad(Y\in \n(\R)).
$$
The map $X\rightarrow \psi_X$ is an isomorphism between $i\overline\n(\R)$ and the unitary characters of $N(\R)$.
It is easy to see $\psi_X$ is non-degenerate if and only if $X\in i\g(\R)$ is a principal nilpotent element.
In this case we write $(B(\R),X)$ for the Whittaker datum $(B(\R),\psi_X)$.

It is clear that $(B(\R),X)\mapsto G(\R)\cdot X$ is a bijection between Whittaker data and
$(\Op(\R)\cap i\g(\R))/G(\R)$. The main result of \cite{matumoto} (Theorem A) says:

@@

\begin{equation}
  \label{e:Epi}
  E_\pi=\sum_{\alpha\in S(\lambda)}X_\alpha\in \N.
\end{equation}

Then $E_\pi\in\s$ is a regular nilpotent element, and
$$
\AV(\pi)=\overline{K\cdot E_\pi}.
$$
\end{lemma}

\begin{proof}
If $\pi$ is any discrete series representation, with Harish-Chandra parameter $\lambda$, let
$\n_{\lambda}=\sum_{\alpha\in\Delta^+(\lambda)}\g_\alpha$.
Then by \cite[Proposition 6.8]{vogan_irreducibility} $AV(\pi)=K\cdot(\n_\lambda\cap\s)$.

Now assume $\pi$ is large. This implies $X_\alpha\in\s$ for all $\alpha\in S(\lambda)$, so $E_\pi\in \s$. 
By~\cite{kostant_tds} $E_\pi$ is a regular nilpotent element, so the closure of $K\cdot E_\pi$ in $\s$ is equal to $K\cdot(\n_\lambda\cap\s)$. 
\end{proof}

@@

Suppose $\pi$ is a discrete series representation.
Choose a compact Cartan subgroup $T$, and let $\lambda\in\t^*$ be the Harish-Chandra parameter of $\pi$.
Identify $\lambda$ with an element $H_\pi\in i\t\subset i\g$ using the isomorphism $\t\simeq \t^*$ induced by the Killing form. 
Define
\begin{equation} \label{semisimple_orbit_HC} \mathcal{O}_{\pi}=\quad \text{$G(\R)$-orbit of the Harish-Chandra parameter $H_\pi$}.\end{equation}

Consider the \emph{asymptotic cone} of $\mathcal{O}_\pi$ (see e.g. \cite[Section 3]{AVAV}) :
\[ \AC(\mathcal{O}_\pi) = \left\{ v \in \g \ : \ \text{$\exists (\varepsilon_n, x_n)\in (\R_+ \times \mathcal{O}_\pi)^\mathbb{N}$ such that $\varepsilon_n \to 0$ and $\varepsilon_n x_n \to v$}  \right\}. \]

@@

We recall a recent result of the first author (JDA) and Kaletha, which spells out the connection between large discrete series and Whittaker data using  {\it Kostant sections}. Let~$X$ be a regular nilpotent element of~$\g$, and suppose~$X$ fits into an $\SL(2)$-triple $(X, H, Y)$. The Kostant section~$\Kostant{X}$ is the affine subspace $X + \mathrm{Cent}_{\g}(Y)$ of~$\g$. It depends on the choice of $\SL(2)$-triple, but if $X \in \g(\R)$, then the $G(\R)$-conjugacy class of~$\Kostant{X}$ depends only on the $G(\R)$-conjugacy class of~$X$.

\begin{proposition}[\cite{adams_kaletha}]\label{JeffTasho_criterion}
Suppose $\pi$ is a generic discrete series representation of~$G(\R)$. Suppose~$\pi$ be a Whittaker datum for~$G(\R)$, and let~$X \in i\g(\R)$ be a regular nilpotent element such that $\w = \w_X$. 
Then $\pi$ is $\w$-generic if and only if $\Kostant{X}$ meets $\mathcal{O}_\pi$. 
\end{proposition}

@@

Theorem~\ref{th:main} gives a bijection between the set $(\Op \cap \s)/K$ of principal nilpotent~$K$-orbits and the set $(\Op \cap i\g(\R))/G(\R)$ of principal nilpotent~$G(\R)$-orbits.

@@

Suppose $E_\R\in \Op\cap i\g(\R)$. Then we may choose an $SL(2)$-triple $(H_\R,E_\R,F_\R)$ with $F_\R$  contained in $i\g(\R)$,
which implies $H_\R\in \g(\R)$. 
After conjugating by~$G(\R)$ we may further assume $\theta(E_\R)=-F_\R$.
The Sekiguchi correspondence takes the $G(\R)$-orbit of~$E_\R$ to the $K$-orbit of
$$
E_\theta=\frac12(E_\R-F_\R+H_\R)\in \N\cap \s.
$$

Computing the map in the other direction goes as follows.
Suppose $E_\theta\in \Op\cap\s$. Choose an $SL(2)$-triple  $(E_\theta, H_\theta,F_\theta)$ with
$F_\theta\in\s$, which implies $H_\theta\in\k$. After conjugating by $K$ we may further assume $\sigma(E_\theta)=F_\theta$.
Then the Sekiguchi correspondence takes the $K$-orbit of $E_\theta$ to the $G(\R)$-orbit of 
\begin{equation}\label{def_E_R}
E_\R=\frac 12(E_\theta-F_\theta-H_\theta)\in\N\cap i\g(\R).
\end{equation}

@@

Consider the principal nilpotent element~$E_\pi$ defined in~\eqref{e:Epi}. We define another regular nilpotent element  $E_\R$ by using $E_\pi$ as $E_\theta$ in~\eqref{def_E_R}. We know that the associated variety of $\pi$ is the closure of $K \cdot E_\pi$ (Lemma~\ref{l:pi_to_av}), so the image of $\AV(\pi)$ by the Sekiguchi correspondence is the closure of the $G(\R)$-orbit of $E_\R$. We want to check that $\WF(\pi)$

By Lemma~\ref{lem:matumoto} it is enough to check that $\pi$ is $\w$-generic for the Whittaker datum $\w$ corresponding to~$E_\R$. By Proposition~\ref{JeffTasho_criterion} this is the case if and only if the Kostant section $\Kostant{E_\R}$ meets $\mathcal{O}_\pi$.

Now, we know that $\Kostant{E_\R}=E_\R+\mathrm{Cent}_\g(F_\R)$ contains $E_\R-F_\R = H_\theta$\footnote{AA 2024-IV-15: insert the correct version once we've checked the Sekiguchi formulas in the previous paragraph. }.
In general the orbits $\mathcal{O}_\pi$ and $G(\R) \cdot H_\theta$ are disjoint. 
Nevertheless $H_{\lambda}$ and $H_\theta=[E_\theta, F_\theta]$ are in the same Weyl chamber\footnote{AA 2024-IV-15: should probably explain a bit more why this is correct.}.
Therefore the result follows from:

@@

\begin{lemma}\label{lem:Kost_and_AC} Let $\mathcal{O}$ be a regular semisimple orbit and  let $X$ be a regular nilpotent element. Then $\Kostant{X}$ meets~$\mathcal{O}$ if and only if $X$ is in the asymptotic cone~$\AC(\mathcal{O})$. 
\end{lemma}
\begin{proof} See if we spell it  out (one direction is in Adams--Kaletha, but parts of the arguments are similar so it might be more convenient include both directions...). Maybe it's actually possible just refer to Fresse--Mehdi with pointers. \end{proof}

@@

\begin{lemma}
  Let $\pi$ be a large discrete series representation.
Choose a compact Cartan subgroup $T$ and let $\lambda\in \t^*$ be the Harish-Chandra parameter of $\pi$.
Define $E_\theta=E_\pi$ as in \eqref{e:Epi}, and choose an $SL(2)$-triple $(H_\theta,E_\theta,F_\theta)$ as usual.
After conjugating by $K$ we may assume $\sigma(E_\theta)=F_\theta$.  Then\footnote{AA 14-II-2023: no closure?}
$$
\WF(\pi)=G(\R)\cdot\frac i2(E_\theta-F_\theta-H_\theta)\in\N\cap i\g(\R)
$$
Conversely if $E_\R\in \Op\cap i\g(\R)$, choose an $SL(2)$-triple $(H_\R,E_\R,F_\R)$ which (after conjugating by $G(\R)$) satisfies $\theta(E_\R)=-F_\R$.
Let
$$
H_\theta=E_\R-F_\R.
$$
Then $H_\theta\in\k$ is a regular semisimple element.
Set $\t=\Cent_{\g}(H_\theta)$, and let
$$
\Delta^+=\{\alpha\in\Delta(T,G)\mid  \alpha(H_\theta)>0\}
$$
Then $\pi$ is the discrete series representation in $\Pi$ whose Harish-Chandra parameter is dominant for $\Delta^+$. 
\end{lemma}

\subsection{Bonus}

>>>>

\subsubsection*{Nilpotent orbits} 
\label{s:nilpotentorbits}
Let $\N$ be the set of nilpotent elements of $\g$.

We need to pass to the dual Lie algebra $\g^*$. We follow  \cite[Section 2]{AVAV}.
Define $\N^*\subset \g^*$ to be the set of elements whose orbit is a cone.
We identify $\g$ and $\g^*$ using  a non-degenerate invariant form on $\g$.
This restricts to an isomorphism $\N\simeq \N^*$. Although this isomorphism depends on the choice of form,
the bijection on the level of $G$-orbits is canonical.

We define $\O_p$ to be the principal nilpotent orbit in $\N^*$: this is the unique orbit such that $\overline{\O_p}=\N^*$.

In the $\theta$-picture we have an analogous construction. We may assume the form is orthogonal with respect to $\theta$.
Consequently the decomposition $\g=\mathfrak k+\s$ also gives $\g^*=\mathfrak k^*\oplus \s^*$, and this induces
a canonical bijection between $K$-orbits on $\s$ and on $\s^*$.

@@

Suppose $\pi$ is an irreducible $(\g,K)$-module.
Let $I_\pi$ be the annihlator of $\pi$ in the universal enveloping algebra.
Let $\AVann(\pi)\subset\N^*\subset \g^*$ be the associated variety of $I_\pi$.
This is the closure of a single complex nilpotent orbit $\O$
See \cite[Section 1, Theorem 4.7]{vogan_bowdoin}.

The associated variety of $\pi$, denoted $\AV(\pi)$, is a closed, $K$-invariant set in $\s^*$, so it is a finite union of
nilpotent $K$-orbits on $\s^*$. See \cite[Section 5]{vogan_bowdoin}.

@@

We define the associated variety of a Hilbert space representations to be that of  its underlying $(\g,K)$-module.

@@

We define the wave front set of a $(\g,K)$-module to be that of its 
Hilbert space globalization.

@@

We write $\Wh(\pi)$ for the set of Whittaker data~$\w$ such that $\pi$ has a $\w$-Whittaker model.

We define the Whittaker data  of a $(\g,K)$-module to be that of its 
Hilbert space globalization.

@@

\begin{proposition} \label{invariants_ds}
  Suppose $\pi$ is a generic discrete series representation of $G(\R)$. Then:
\begin{enumerate}[(a)]
\item $\Wh(\pi)$ is a single Whittaker datum;
\item $\AV(\pi)$ is the closure of a single $K$-orbit on~$\mathcal{O}_p \cap \s^*$;
\item  $\WF(\pi)$ is the closure of a single $G(\R)$-orbit on $\mathcal{O}_p \cap  i \g(\R)^*$.
\end{enumerate}
\end{proposition}

\begin{proof} For (a), by \cite{vogan-gelfand-kirillov} and \cite{kostant_whittaker}
a representation $\pi$ is large if and only if it admits a Whittaker model for some Whittaker datum,
and by \cite[Lemma 14.14]{abv} a  large discrete series representation admits a unique Whittaker datum.
For (b) see \cite[Proposition A.9]{AV1}.

@@

Because of this result we abuse notation slightly as follows. If~$\pi$ is a generic discrete series representation of~$G(\R)$, we write $AV(\pi)=\O$ to indicate that $AV(\pi)$ is the closure of the $K$-orbit $\O\subset \s^*$.
Similarly we write $\WF(\pi)=\O$ to indicate that $\WF(\pi)$ is the closure of the $G(\R)$-orbit $\O\subset i\g(\R)^*$.

@@

\begin{lemma}
  \label{l:large}
  The following conditions are equivalent.
  \begin{enumerate}
    \item $\pi$ is large;
\item $\AVann(\pi)=\N^*=\overline{\Op}$;
\item $\dim(\AVann(\pi))=N$;
  \item $\dim(\AV(\pi))=N/2$;
  \item $\GK(\pi)=N/2$;
    \item $\pi$ is generic.

@@

Given a generic discrete series representation~$\pi$ we consider the invariants of~$\pi$ defined by Proposition~\ref{invariants_ds}: a Whittaker datum  $\Wh(\pi)$, a $K$-orbit $\AV(\pi)$ in $\mathcal{O}_p \cap \s^*$, and a $G(\R)$-orbit $\WF(\pi)$ in $\mathcal{O}_p \cap  i \g(\R)^*$.

\begin{theorem} \label{th:main} Suppose~$\Pi$ is an $L$-packet of discrete series for~$G(\R)$. 
The maps $\pi \mapsto \Wh(\pi)$, $\pi \mapsto \AV(\pi)$ and $\pi\mapsto \WF(\pi)$ induce bijections between:
\begin{enumerate}
\item[(1)] The set~$\Pi_{\mathrm{gen}}$ of generic discrete series representations in~$\Pi$ ;
\item[(2)] The set of Whittaker data for $G(\R)$ ;
\item[(3)] The set of~$K$-orbits on $\mathcal{O}_p \cap \s^*$.
\item[(4)] The set of~$G(\R)$-orbits on $\mathcal{O}_p \cap  i \g(\R)^*$.
\end{enumerate}
In particular if $\pi$ is a large discrete series represenation then each of these three invariants of $\pi$ determines the other $2$. 
\end{theorem}

Before embarking on the proof, let us point out that the maps in the Theorem are defined using rather deep results of representation theory. However, we shall show in Section~\ref{sec:explicit} that all bijections $(i) \leftrightarrow (j)$, for $i,j \in \{(1), \dots, (4)\}$, can be spelled out in elementary terms. Furthermore we shall prove, using elementary arguments, that (3) $\leftrightarrow$ (4) coincides with the Kostant--Sekiguchi correspondence.

@@

Note that $Q_\sigma(G)$ is a group of automorphisms of $G(\R)$. It therefore acts on representations of $G(\R)$. 
Similarly $Q_\theta(G)$ is a group of automorphisms of $(\g,K)$, and acts on $(\g,K)$-modules.\footnote{AA 30-IV-2024: shouldn't this be:  $\Gad^\sigma$ and $\Gad^\theta$  act on $G(\R)$-representations or $(\g,K)$-modules respectively, and this induces actions of $Q_{\sigma}(G)$ and $Q_{\theta}(G)$ on equivalence classes...? JDA this is the same thing}
These two actions are intertwined by the isomorphism of Lemma~\ref{lem:iso_q}:

@@

To define the action of~$Q_{\sigma}(G)$ on~(4), first observe that~$\Gad$ acts on~$\g$ by the adjoint action. This action preserves~$\Op$, and its restriction to~$\Gad(\R)$ preserves~$\Op\cap i\g(\R)^*$. Since elements on a given~$G(\R)_{\mathrm{ad}}$-orbit are in the same~$G(\R)$-orbit, this induces an action of $Q_{\sigma}(G)=\Gad(\R)/G(\R)_{\mathrm{ad}}$ on~$(\Op\cap i\g(\R)^*)/G(\R)$.

\begin{proposition}\label{prop:action_on_real_orbits}
\begin{enumerate} 
\item The action of~$Q_{\sigma}(G)$ on $(\Op\cap i\g(\R)^*)/G(\R)$ is simply transitive.
\item The map $\pi \mapsto \WF(\pi)$ is $Q_{\sigma}(G)$-equivariant.
\end{enumerate}
\end{proposition}

\begin{proof}
The proof of~(1) is a standard argument in group cohomology. Suppose $\omega,\omega'$ are $G(\R)$ orbits in $\Op\cap i\g(\R)^*$,
and choose $X\in \omega,X'\in\omega'$. Then there exists $g\in G$ such that $\Ad(g)(X)=X'$. Since $X\in \g(\R)^*$ the condition $X'\in i\g(\R)^*$ 
is $g\inv \sigma(g)\in \Stab_G(X)$.  Since $X$ is principal $\Stab_G(X)=ZU$ where $U$ is a unipotent group. 
A standard fact is that $H^1(\sigma, U)=1$ \cite[Chap.~III, Proposition~6]{Serre_Galois}, and from this we see that, after multiplying $g$ on the right by an element of $u$, we can assume $g\inv \sigma(g)\in Z$. This is equivalent to: the image of $g$ in $\Gad$ is in $\Gad(\R)$. On the other hand $\omega=\omega'$ 
if and only if $X'=\Ad(g)X$ for some $g\in G(\R)$. This competes the proof of (1).

@@

On the other hand $Q(G)$, realized as $Q_\theta(G)$ is a group of automorphisms of $(\g,K)$.
It induces actions on $\s^*$, $\Op$ and $\Op\cap\s^*$, and on $(\g,K)$-modules.

\begin{proposition}\label{prop:action_on_K_orbits}
\begin{enumerate} 
\item The action of~$Q(G)$ on $(\Op \cap \s^*)/K$ is simply transitive.
\item The map $\pi \mapsto \AV(\pi)$ is $Q(G)$-equivariant.
\end{enumerate}
\end{proposition}

@@

First we describe the set of Whittaker data more explicitly.
Let $B$ be a Borel subgroup of $G$. Let $N$ be the nilradical of $B$, and $\overline N$ the opposite
nilpotent subgroup. Let $\n,\overline\n$ be the Lie algebras of $N,\overline N$.
If $B$ is defined over $\R$ we consider $N(\R),\n(\R), \overline\n(\R)$, etc.
For $X\in i\overline \n(\R)^*$ define a unitary character $\psi_X$ of $N(\R)$ by:
$$
\psi_X(e^Y)=e^{X(Y)}\quad(Y\in \n(\R)).
$$
The map $X\rightarrow \psi_X$ is an isomorphism between $i\overline\n(\R)^*$ and the unitary characters of $N(\R)$.
It is easy to see $\psi_X$ is non-degenerate if and only if $X\in i\g(\R)^*$ is a principal nilpotent element.
In this case we write $(B(\R),X)$ for the Whittaker datum $(B(\R),\psi_X)$.

It is clear that $(B(\R),X)\mapsto G(\R)\cdot X$ is a bijection between Whittaker data and
$(\Op(\R)\cap i\g(\R)^*)/G(\R)$. The main result of \cite{matumoto} (Theorem A) says:

@@

\begin{equation}
  \label{e:Fpi}
  F_\pi=\sum_{\alpha\in S(\lambda)}X_{-\alpha}\in \N.
\end{equation}

Then $F_\pi\in\s$ is a regular nilpotent element. Let $\xi_\pi\in \N^*$ be the corresponding element (cf. Section \ref{s:nilpotentorbits}).
Then 
$$
\AV(\pi)=\overline{K\cdot \xi_\pi}.
$$
\end{lemma}

See
\cite[Propositions A.7 and A.9]{AV1}.

@@

Suppose $\pi$ is a discrete series representation.
Choose a compact Cartan subgroup $T$, and let $\lambda\in\t^*$ be the Harish-Chandra parameter of $\pi$.
Define
\begin{equation} \label{semisimple_orbit_HC} \mathcal{O}_{\pi}=G(\R)\cdot\lambda.\end{equation}

Consider the \emph{asymptotic cone} of $\mathcal{O}_\pi$ (see e.g. \cite[Section 3]{AVAV}) :
\[ \AC(\mathcal{O}_\pi) = \left\{ v \in \g^* \ : \ \text{$\exists (\varepsilon_n, x_n)\in (\R_+ \times \mathcal{O}_\pi)^\mathbb{N}$ such that $\varepsilon_n \to 0$ and $\varepsilon_n x_n \to v$}\}  \right\}. \]
This is a union of nilpotent $G(\R)$-orbits on $i\g(\R)^*$.

@@

We recall a recent result of the first author and Tasho Kaletha, which spells out the connection between large discrete series and Whittaker data using  {\it Kostant sections}. Let~$X$ be a regular nilpotent element of~$\g$, and choose an  $\SL(2)$-triple $(X, H, Y)$. The Kostant section~$\Kostant{X}$ is the affine subspace $X + \mathrm{Cent}_{\g}(Y)$ of~$\g$. This depends on the choice of $\SL(2)$-triple, but if $X \in \g(\R)^*$, then the $G(\R)$-conjugacy class of~$\Kostant{X}$ depends only on the $G(\R)$-conjugacy class of~$X$.

\begin{proposition}[\cite{adams_kaletha}]\label{JeffTasho_criterion}
Suppose $\pi$ is a generic discrete series representation of~$G(\R)$. Suppose~$\pi$ be a Whittaker datum for~$G(\R)$, and let~$X \in i\g(\R)^*$ be a regular nilpotent element such that $\w = \w_X$. 
Then $\pi$ is $\w$-generic if and only if $\Kostant{X}$ meets $\mathcal{O}_\pi$. 
\end{proposition}

@@

Theorem~\ref{th:main} gives a bijection between the set $(\Op \cap \s^*)/K$ of principal nilpotent~$K$-orbits and the set $(\Op \cap i\g(\R)^*/G(\R)$ of principal nilpotent~$G(\R)$-orbits.

@@

{\bf Note:} Need some $*$'s here.

Suppose $F_\R\in \Op\cap i\g(\R)$. Then we may choose an $SL(2)$-triple $(H_\R,E_\R,F_\R)$ with $E_\R$  contained in $i\g(\R)$,
which implies $H_\R\in \g(\R)$. 
After conjugating by~$G(\R)$ we may further assume $\theta(E_\R)=-F_\R$.
The Sekiguchi correspondence takes the $G(\R)$-orbit of~$E_\R$ to the $K$-orbit of
$$
F_\theta=\frac12(-E_\R+F_\R+H_\R)\in \N\cap \s.
$$

Computing the map in the other direction goes as follows.
Suppose $F_\theta\in \Op\cap\s$. Choose an $SL(2)$-triple  $(E_\theta, H_\theta,F_\theta)$ with
$E_\theta\in\s$, which implies $H_\theta\in\k$. After conjugating by $K$ we may further assume $\sigma(E_\theta)=F_\theta$.
Then the Sekiguchi correspondence takes the $K$-orbit of $E_\theta$ to the $G(\R)$-orbit of 
\begin{equation}\label{def_F_R}
F_\R=\frac 12(-E_\theta+F_\theta-H_\theta)\in\N\cap i\g(\R).
\end{equation}

@@

Consider the principal nilpotent element~$F_\pi$ defined in~\eqref{e:Fpi}. We define another regular nilpotent element  $F_\R$ using 
\eqref{def_F_R} using $F_\theta=F_\pi$ on the right hand side. 
We know that  $\AV(\pi)=K \cdot \xi_\pi$ (where $\xi_\pi\in\g^*$ corresponds to $F_\pi$, see Lemma~\ref{l:pi_to_av}), so the image of $\AV(\pi)$ by the Sekiguchi correspondence is 
 $G(\R)\cdot F_\R$. We want to check that $\WF(\pi)$

By Lemma~\ref{lem:matumoto} it is enough to check that $\pi$ is $\w$-generic for the Whittaker datum $\w$ corresponding to~$F_\R$. By Proposition~\ref{JeffTasho_criterion} this is the case if and only if the Kostant section $\Kostant{F_\R}$ meets $\mathcal{O}_\pi$.

Now, we know that $\Kostant{F_\R}=F_\R+\mathrm{Cent}_\g(F_\R)$ contains $F_\R-F_\R = H_\theta$\footnote{AA 2024-IV-15: insert the correct version once we've checked the Sekiguchi formulas in the previous paragraph. }.
In general the orbits $\mathcal{O}_\pi$ and $G(\R) \cdot H_\theta$ are disjoint. 
Nevertheless $H_{\lambda}$ and $H_\theta=[F_\theta, F_\theta]$ are in the same Weyl chamber\footnote{AA 2024-IV-15: should probably explain a bit more why this is correct.}.
Therefore the result follows from:

@@

\begin{lemma}\label{lem:Kost_and_AC} Let $\mathcal{O}$ be a regular semisimple orbit and  let $X$ be a regular nilpotent element. Then $\Kostant{X}$ meets~$\mathcal{O}$ if and only if $X$ is in the asymptotic cone~$\AC(\mathcal{O})$.

Consequently each of the three invariants $\Wh(\pi),\AV(\pi)$ and $\WF(\pi)$ are constant for $\pi=\pi(\lambda)$ a large discrete series representation, and $\lambda$ in a fixed large Weyl chamber.
\end{lemma}
\begin{proof} 
If $X\in \AC(\mathcal O)$ then $\Kostant{X}$ meets $\mathcal O$ by \cite[Lemma 3.2.3]{adams_kaletha}.

Suppose $\Kostant{X}$ meets $\mathcal O$.
This means there is an $SL(2)$-triple $(H,Y,X)$, and $Z\in \Cent(Y)$ such that $X+Z\in \O$. 
By the representation theory of $SL(2)$ we know that $Z=\sum_{j=0}^n Z_j$ with $[H,Z_j]=-jZ_j$. 
Then 
$$
\Ad(e^{tH})(X+Z)=e^{2t}(X) + \sum_j e^{-tj}Z_j
$$
Taking the limit as $t\rightarrow 0$ gives $\Ad(e^{tH})(X+Z)\rightarrow e^{2t}X$. Therefore
$$
\lim_{t\rightarrow 0}e^{-2t}\Ad(e^{tH})(X+Z)=X
$$ which implies the result.
\end{proof}

For a closely related result see \cite[Proposition 3.5]{fm}.

@@

\begin{lemma}
  Let $\pi$ be a large discrete series representation.
Choose a compact Cartan subgroup $T$ and let $\lambda\in \t^*$ be the Harish-Chandra parameter of $\pi$.
Define $F_\theta=F_\pi$ as in \eqref{e:Fpi}, and choose an $SL(2)$-triple $(H_\theta,E_\theta,F_\theta)$ as usual.
After conjugating by $K$ we may assume $\sigma(E_\theta)=F_\theta$.  Then\footnote{AA 14-II-2023: no closure?}
$$
\WF(\pi)=G(\R)\cdot\frac i2(E_\theta-F_\theta-H_\theta)\in\N\cap i\g(\R)
$$
Conversely if $E_\R\in \Op\cap i\g(\R)$, choose an $SL(2)$-triple $(H_\R,E_\R,F_\R)$ which (after conjugating by $G(\R)$) satisfies $\theta(E_\R)=-F_\R$.
Let
$$
H_\theta=E_\R-F_\R.
$$
Then $H_\theta\in\k$ is a regular semisimple element.
Set $\t=\Cent_{\g}(H_\theta)$, and let
$$
\Delta^+=\{\alpha\in\Delta(T,G)\mid  \alpha(H_\theta)>0\}
$$
Then $\pi$ is the discrete series representation in $\Pi$ whose Harish-Chandra parameter is dominant for $\Delta^+$. 
\end{lemma}

Here is another application of Proposition \ref{prop:AC_chamber}.

\begin{lemma}
\label{l:equalWhittaker}
Suppose $\pi=\pi(\lambda),\pi'=\pi(\lambda')$ are large discrete series
representations. Then $\Wh(\pi)=\Wh(\pi')$ if and only if
$\lambda,\lambda'$ are in the same Weyl chamber (up to the action of
$W_K$).
\end{lemma}

\begin{proof}
If $\lambda,\lambda'$ are in the same Weyl chamber the $\Wh(\pi)=\WF(\pi')$ by Lemma \ref{lem:Kost_and_AC}.
